---
title: "Sequencing report"
subtitle: "QC and basic stats"
author: "Daniel Fischer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      toc_collapsed: true
    number_sections: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
.libPaths(c("/projappl/project_2001746/R/libraries", .libPaths()))
#library("knitr")
#library("GenomicTools")
#library("Luke")
library("xtable")   # Needed for LaTeX output of the tables
#library("viridis")  # Needed for the colouring of the plots
library("rjson")    # Needed for multiqc dgsb etail data
#library("adegenet")
#library("vcfR")
library("DT")
library("kableExtra")
options(scipen=999,
        stringsAsFactors=FALSE)
knitr::opts_chunk$set(echo = FALSE,
                      cache = FALSE,
                      cache.lazy = FALSE,
         #             dev = c('png', 'pdf'),
                      fig.align = 'center', fig.height = 5, fig.width = 8.5)
if(!is.element("snakemake",ls())){
  projFolder <- "/scratch/project_2001881/201214_NB551722_0019_AHYGCTBGXC"
  pipelineFolder <- "/scratch/project_2001881/Pipeline-SequencingQC/"
  reportFolder <- "/scratch/project_2001881/201214_NB551722_0019_AHYGCTBGXC_QC-report"
  expected_reads_per_sample <- 100000
}
noOfLanes <- 4
```

```{r import reads per sample}
no_reads_per_file.files <- list.files(file.path(reportFolder,"data"))

no_reads_per_file <- list()

for(i in 1:length(no_reads_per_file.files)){
  no_reads_per_file[[i]] <- read.table(file.path(reportFolder,"data",no_reads_per_file.files[i]))
}
```

```{r filter no read per sample table}
no_reads_per_file.R1 <- list()
for(i in 1:length(no_reads_per_file.files)){
  no_reads_per_file.R1[[i]] <- no_reads_per_file[[i]][grep("R1_001",no_reads_per_file[[i]][,2]),]
}
```

# Basic statistics
```{r}
readFilesR1 <- list.files(file.path(reportFolder,"data"))
subProjects <- sapply(strsplit(readFilesR1,"\\."),"[",1)

# Lanes per subproject
lanesPerSubproject <- c()
for(i in 1:length(subProjects)){
  lanesPerSubproject[i] <- length(grep("R1_001.fastq.gz",no_reads_per_file[[i]][,2]))
}

# Samples per subproject (THIS IS VERY, VERY BAD CODED AND WORKS ONLY IS _S IS NOT PART OF THE SUBGROUP NAME!!!! FIX THAT WHEN THERE IS MORE TIME!!!)
samplenamesPerSubproject <- list()
samplesPerSubproject <- c()
for(i in 1:length(subProjects)){
  samplenamesPerSubproject[[i]] <- unique(sapply(strsplit(no_reads_per_file[[i]][,2], "_S"),"[",1))
  samplesPerSubproject[i] <- length(samplenamesPerSubproject[[i]])
}
```


```{r no of reads per sample}
no_reads_per_sample.R1 <- list()
for(i in 1:length(subProjects)){
    no_reads_per_sample.R1[[i]] <- as.data.frame(matrix(0, ncol=2, nrow=length(samplenamesPerSubproject[[i]])))
}

for(i in 1:length(subProjects)){
  for(j in 1:length(samplenamesPerSubproject[[i]])){
    no_reads_per_sample.R1[[i]][j,2] <- samplenamesPerSubproject[[i]][j]
    no_reads_per_sample.R1[[i]][j,1] <- sum(no_reads_per_file.R1[[i]][grep(samplenamesPerSubproject[[i]][j],no_reads_per_file.R1[[i]][,2]),1])
  }
}
```

## Considered Subprojects
There are `r length(subProjects)` different projects inside the sequencing run, namely:

```{r report basic stats}
out <- data.frame(c("Subprojects",
                    "Lanes per subproject",
                    "Samples per subproject"),
                  c(subProjects,
                    lanesPerSubproject,
                    samplesPerSubproject))

out_html <- knitr::kable(out, col.names = NULL, "html")
kable_styling(out_html, "striped", position = "left")
```

## Deviation from expected No. of reads
Expected reads (for paired-end, reads refer to fragments, meaning that the both reads R1 and R2 form one read) were `r expected_reads_per_sample`

```{r deviation from expectation}
deviationR1 <- list()
for(i in 1:length(no_reads_per_file.files)){
  deviationR1[[i]] <- no_reads_per_file.R1[[i]][,1]/4-(expected_reads_per_sample/noOfLanes)
  tmpCol <- sign(deviationR1[[i]])
  tmpCol[tmpCol==-1] <- "darkred"
  tmpCol[tmpCol==1] <- "darkgreen"
  barplot(deviationR1[[i]], col=tmpCol, border=NA)
}
```

A list of lanes with less reads than expected (for read R1)

```{r table reads per lane,results = 'asis'}
i <- 1
#for(i in 1:length(no_reads_per_file.files)){
  tmp <- no_reads_per_file.R1[[i]][(no_reads_per_file.R1[[i]][,1]/4)<expected_reads_per_sample/noOfLanes,]
  if(nrow(tmp)>0){
    tmp[,1] <- tmp[,1]/4
    colnames(tmp) <- c("No reads", "Lane")
    datatable(tmp)
#    knitr::kable(tmp) %>% kable_styling
#  }
}
```

## Deviation from expected No. of reads (sample-wise)

```{r deviation from expectation per sample}
deviationR1.samples <- list()
for(i in 1:length(no_reads_per_sample.R1)){
  deviationR1.samples[[i]] <- no_reads_per_sample.R1[[i]][,1]/4-(expected_reads_per_sample)
  tmpCol <- sign(deviationR1.samples[[i]])
  tmpCol[tmpCol==-1] <- "darkred"
  tmpCol[tmpCol==1] <- "darkgreen"
  barplot(deviationR1.samples[[i]], col=tmpCol, border=NA)
}
```

```{r table reads per sample,results = 'asis'}
i <- 1
#for(i in 1:length(no_reads_per_file.files)){
  tmp <- no_reads_per_sample.R1[[i]][(no_reads_per_sample.R1[[i]][,1]/4)<expected_reads_per_sample,]
  if(nrow(tmp)>0){
  # Fixing that we count lines, for fastq we have to divide by 4 then instead
    tmp[,1] <- tmp[,1]/4
    colnames(tmp) <- c("No reads", "Sample")
    datatable(tmp)
#    knitr::kable(tmp) %>% kable_styling
#  }
}
```

## Reads per sample
This table contains all samples and their number of reads

```{r table reads all sample,results = 'asis'}
i <- 1
#for(i in 1:length(no_reads_per_file.files)){
  tmp <- no_reads_per_sample.R1[[i]]
  if(nrow(tmp)>0){
  # Fixing that we count lines, for fastq we have to divide by 4 then instead
    tmp[,1] <- tmp[,1]/4
    colnames(tmp) <- c("No reads", "Sample")
    datatable(tmp)
#    knitr::kable(tmp) %>% kable_styling
#  }
}
```